---
title: "Viz 1 & 2_EDA"
author: "Jana Lee"
date: "9/26/2019"
output: html_document
---
# viz_and_eda

This is the repo for examples related to Viz and EDA.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Load in tidyverse and ggridges.
```{r}
library(tidyverse)
library(ggridges)
```

Loading in rnoaa package, the r package that loads in national weather service data. They make data publicly available. This will be the dataset we are going to use
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## create a ggplot
The follow below happens because we definied the dataset itself and the x and y axes. However, we haven't told it what plot we are going to make yet.
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))

# adding geompoint allows us to plot this.
ggplot(weather_df, aes(x = tmin, y = tmax)) + geom_point()
```

An alternate way of making this plot. Creating a plot and printing it in the same step.
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Saving Initial plots.
```{r}
scatterplot = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

# this is useful when you want to create a plot and you want to save it into an initial folder for people to reference.
scatterplot
```

Adding color.
alpha blending takes a value that's not completely transparent. Get a little bit of transparency.
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name, alpha = .40))
```

Why do aes positions matter?
Turn of SE (standard errors - False)
first:
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name, alpha = .40)) +
  geom_smooth (se = FALSE)
```

vs. 

second
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(aes(color = name, alpha = .40)) +
  geom_smooth (se = FALSE)
```

Facets
This produces 3 different graphs with name in the columns! Separated out by 3 distinct panels. Which of these is better is a matter of judgment.
- harder to do comparisons between each of the curves
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

But now I want to see the seasonal trend of precipitation! Geom_point = aes makes big dots.
```{r}
weather_df %>%
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(size = 2, se = FALSE)

```

## some extra stuff
You can remove the data points by eliminating the geom points. You can pick and choose what you want to show. Geom that we pick are up to us!
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE)
```

2d Density
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_bin2d()

```

## Univariate plots
Creating a histogram for the x variable of max temperatures.You can add fill to fill in the bars, as compared to color = name.
```{r}
weather_df %>%
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram()
```

If we don't like stacked histograms, we can use position = dodge.
```{r}
weather_df %>%
  ggplot(aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
```

Seeing surface area under curve. Pretty!
```{r}
weather_df %>%
  ggplot(aes(x = tmax, fill = name)) + 
  geom_density(alpha = .3)
```

Boxplots
```{r}
weather_df %>%
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot(alpha = .3)
```

Violin plots
```{r}
weather_df %>%
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()
```

```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  stat_summary(fun.y = median, geom = "point", color = "blue", size = 4)
```

Ridge plots
```{r}
weather_df %>%
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges(scale = .85)
```

## Saving a plot. Make sure you write the code, bc exporting it will not work.
```{r}
weather_plot = ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 

ggsave("weather_plot.pdf", weather_plot, width = 8, height = 5)

```

Can only be as wide as the document is
```{r, fig. width = 12}
weather_df %>%
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(alpha = .4) +
  geom_smooth(se = FALSE)
```


# Lecture 2: Viz and EDA Part 2

## making new plots

```{r}
weather_df %>%
    ggplot(aes(x = tmin, y = tmax, color = name)) + geom_point(alpha = .5)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


## Labels, Ticks
Always label your graphs!
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  )
```

## Scaling 
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, -5, 20),
    labels = c("-15ยบ C", "0 (still too cold", "15"))
# have tick marks wherever you want it! It is under breaks.
```



```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ยบC", "0", "15"),
    limits = c(-20, 30)) + 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right")
# This might drop a few rows bc can't take sqrt of negative (90 rows containing missing values removed)

# You can also change scale_y (sqrt) to change the scale for y.
```

## Colors & Themes
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_color_hue(name = "Location", h = c(100, 200))

# Choosing colors by hand will make you go crazy, so don't spend so much time on this.
```

## Viridis package
Solves a very specific problem in tidyverse.We downloaded this in Problem 0. As a general rule, use this color palette.

```{r}
ggp_temp_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )
# In place of scale_color_hue, we're going to use viridis here. We have to specifity that discrete = TRUE!
```

## Themes 
Global things that describe what the plot looks like, the behavior of the plot.
Legend Position: 

```{r}
ggp_temp_plot + 
  theme(legend.position = "bottom")

# Put legend on the bottom preserves more space for the actual plot itself.

# The default of the background is gray. We can change this.

ggp_temp_plot + 
  theme_bw() + 
  theme(legend.position = "bottom")

ggp_base + 
  theme_minimal () 
  theme(legend.position = "bottom")

ggp_temp_plot + 
  theme_classic() + 
  theme(legend.position = "bottom")
```

How sensitive should we be about the warning messages? Moderately... generating bc missing values? Nothing we can do about that



