---
title: "Viz 1_EDA"
author: "Jana Lee"
date: "9/26/2019"
output: html_document
---
# viz_and_eda

This is the repo for examples related to Viz and EDA.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Load in tidyverse and ggridges.
```{r}
library(tidyverse)
library(ggridges)
```

Loading in rnoaa package, the r package that loads in national weather service data. They make data publicly available. This will be the dataset we are going to use
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## create a ggplot
The follow below happens because we definied the dataset itself and the x and y axes. However, we haven't told it what plot we are going to make yet.
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))

# adding geompoint allows us to plot this.
ggplot(weather_df, aes(x = tmin, y = tmax)) + geom_point()
```

An alternate way of making this plot. Creating a plot and printing it in the same step.
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Saving Initial plots.
```{r}
scatterplot = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

# this is useful when you want to create a plot and you want to save it into an initial folder for people to reference.
scatterplot
```

Adding color.
alpha blending takes a value that's not completely transparent. Get a little bit of transparency.
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name, alpha = .40))
```

Why do aes positions matter?
Turn of SE (standard errors - False)
first:
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name, alpha = .40)) +
  geom_smooth (se = FALSE)
```

vs. 

second
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(aes(color = name, alpha = .40)) +
  geom_smooth (se = FALSE)
```

